When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-02-12 paper/paper.tex
Correct spacing after et al.

However, subgraph isomorphism algorithms cannot handle \emph{arbitrary} instances involving hundreds
or thousands of vertices. Experimental evaluations of subgraph isomorphism algorithms are usually
performed using randomly generated graph pairs, sometimes together with a mix of real-world graphs,
for example biochemistry and computer vision problems. Using random instances to evaluate
algorithm behaviour can be beneficial, because it provides a way of generating many instances
cheaply, and reduces the risk of over-fitting when tuning design parameters. It also allows
experimenters to control graph parameters such as order, density, or degree distribution, and to
study scaling properties of algorithms with respect to these parameters.  The random instances used
in previous evaluations come from common datasets
\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh) graph
and permuting the vertices. Such instances are guaranteed to be satisfiable---\citet{o:AntonO09}
exploited this property to create large sets of random satisfiable boolean satisfiability instances.
This is the most common approach to generating random subgraph isomorphism instances, meaning
existing benchmark suites contain relatively few non-trivial unsatisfiable instances (although a few
of the patterns in the instances by \citeauthor{DBLP:journals/constraints/ZampelliDS10} have had
extra edges added, making them unsatisfiable). Also, the satisfiable instances tend to be
computationally fairly easy, with most of the difficulty being in dealing with the size of the
model. This has led to bias in algorithm design, to the extent that some proposed techniques, such
as those of \citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.

@@

\citet{DBLP:conf/edbt/KatsarouNT17} present approaches to improve the run times for what they call
``straggler queries''---the small subset of queries that they observe taking much longer than others
to solve.  They observe that permuting graphs (for example, by using degree or label frequency
information) before running the subgraph isomorphism algorithm can improve run times by orders of
magnitude. However, the connection between this approach and variable- and value-ordering heuristics
is not noted, and no consideration is given to dynamic ordering heuristics such as ``smallest domain
first'' \citep{DBLP:journals/ai/HaralickE80}.  Of the orderings proposed, those involving placing
rare labels first are effectively approximations to a static ``smallest domain at top of search
first'' ordering, whilst the remainder use degree as we did earlier in this paper.
\citeauthor{DBLP:conf/edbt/KatsarouNT17} do not investigate any algorithm which employs strong
variable- or value-ordering heuristics, all-different propagation, or even domains, and do not
consider that the apparent successes of their approach could be due to VF2's weaknesses rather than
an inherent property of \NP-completeness.

>>>>

However, subgraph isomorphism algorithms cannot handle \emph{arbitrary} instances involving hundreds
or thousands of vertices. Experimental evaluations of subgraph isomorphism algorithms are usually
performed using randomly generated graph pairs, sometimes together with a mix of real-world graphs,
for example biochemistry and computer vision problems. Using random instances to evaluate
algorithm behaviour can be beneficial, because it provides a way of generating many instances
cheaply, and reduces the risk of over-fitting when tuning design parameters. It also allows
experimenters to control graph parameters such as order, density, or degree distribution, and to
study scaling properties of algorithms with respect to these parameters.  The random instances used
in previous evaluations come from common datasets
\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh) graph
and permuting the vertices. Such instances are guaranteed to be satisfiable---\citet{o:AntonO09}
exploited this property to create large sets of random satisfiable boolean satisfiability instances.
This is the most common approach to generating random subgraph isomorphism instances, meaning
existing benchmark suites contain relatively few non-trivial unsatisfiable instances (although a few
of the patterns in the instances by \citeauthor{DBLP:journals/constraints/ZampelliDS10}\ have had
extra edges added, making them unsatisfiable). Also, the satisfiable instances tend to be
computationally fairly easy, with most of the difficulty being in dealing with the size of the
model. This has led to bias in algorithm design, to the extent that some proposed techniques, such
as those of \citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.

@@

\citet{DBLP:conf/edbt/KatsarouNT17} present approaches to improve the run times for what they call
``straggler queries''---the small subset of queries that they observe taking much longer than others
to solve.  They observe that permuting graphs (for example, by using degree or label frequency
information) before running the subgraph isomorphism algorithm can improve run times by orders of
magnitude. However, the connection between this approach and variable- and value-ordering heuristics
is not noted, and no consideration is given to dynamic ordering heuristics such as ``smallest domain
first'' \citep{DBLP:journals/ai/HaralickE80}.  Of the orderings proposed, those involving placing
rare labels first are effectively approximations to a static ``smallest domain at top of search
first'' ordering, whilst the remainder use degree as we did earlier in this paper.
\citeauthor{DBLP:conf/edbt/KatsarouNT17}\ do not investigate any algorithm which employs strong
variable- or value-ordering heuristics, all-different propagation, or even domains, and do not
consider that the apparent successes of their approach could be due to VF2's weaknesses rather than
an inherent property of \NP-completeness.
